---
output:
  html_document: default
  pdf_document: default
---

<table style="width: 900px"><tr><td bgcolor="darkred"><font color="white" size="10"><div style="text-align: center;"><b>
CANADA COVID-19 Data Analysis
</b></div></font></td></tr></table>

----------------------------------------------------------------------------------------------------------------------------------------------------------------------

<table style="width: 900px"><tr><td bgcolor="darkred"><font color="white" size="6"><b>
&nbsp;&nbsp;1. Introduction
</b></font></td></tr></table>

<table style="width: 900px"><tr><td bgcolor="wheat"><font color="brown" size="5"><b>
&nbsp;&nbsp;1.1 Project Objective
</b></font></td></tr></table>

#### 1. Analyze Canadian COVID-19 data using the dataset from Johns Hopkins University.
#### 2. Compare the outcomes of COVID-19 responses using multiple visualization techniques.
#### 3. Employ cluster analysis to uncover inside relationship among different provinces.

<table style="width: 900px"><tr><td bgcolor="wheat"><font color="brown" size="5"><b>
&nbsp;&nbsp;1.2 Data Source
</b></font></td></tr></table>

#### 1. The data comes from the Johns Hopkins University COVID-19 dataset.
#### 2. JHU's COVID-19 dataset is globally recognized as one of the most comprehensive sources.
#### 3. The dataset includes not only national-level data but also data at the state or provincial level and other detailed breakdowns.

<table style="width: 900px"><tr><td bgcolor="wheat"><font color="brown" size="5"><b>
&nbsp;&nbsp;1.3 Importing Packages
</b></font></td></tr></table>

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(cowplot)
library(ggthemes)
library(forecast)
library(knitr)
library(data.table)
library(leaflet)
library(tinytex)
library(kableExtra)
library(gridExtra)
library(dplyr)
```

<table style="width: 900px"><tr><td bgcolor="wheat"><font color="brown" size="5"><b>
&nbsp;&nbsp;1.4 Importing Dataset
</b></font></td></tr></table>

```{r message=FALSE, warning=FALSE}
url1 <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"
url2 <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"

confirmed <- read_csv(url1)
deaths <- read_csv(url2)
```

<table style="width: 900px"><tr><td bgcolor="darkred"><font color="white" size="6"><b>
&nbsp;&nbsp;2. Data Cleaning and Preprocessing
</b></font></td></tr></table>

<table style="width: 900px"><tr><td bgcolor="wheat"><font color="brown" size="5"><b>
&nbsp;&nbsp;2.1 Checking Data
</b></font></td></tr></table>

```{r message=FALSE, warning=FALSE}
head(confirmed)
```

```{r message=FALSE, warning=FALSE}
head(deaths)
```

```{r message=FALSE, warning=FALSE}
dim(confirmed)
dim(deaths)
```
<table style="width: 900px"><tr><td bgcolor="wheat"><font color="brown" size="5"><b>
&nbsp;&nbsp;2.2 Filtering Observation
</b></font></td></tr></table>

```{r}
confirmed_canada <- confirmed %>% 
  filter(`Country/Region` == "Canada")

deaths_canada <- deaths %>% 
  filter(`Country/Region` == "Canada")
```

<table style="width: 900px"><tr><td bgcolor="wheat"><font color="brown" size="5"><b>
&nbsp;&nbsp;2.3 Remove missing value 
</b></font></td></tr></table>

```{r}
confirmed_cleaned <- confirmed_canada %>%
  filter(!(Long == 0))
deaths_cleaned <- deaths_canada %>%
  filter(!(Long == 0))
```

<table style="width: 900px"><tr><td bgcolor="darkred"><font color="white" size="6"><b>
&nbsp;&nbsp;3. Data Visualization
</b></font></td></tr></table>

<table style="width: 900px"><tr><td bgcolor="wheat"><font color="brown" size="5"><b>
&nbsp;&nbsp;3.1 Total COVID-19 Confirmed Cases by Province
</b></font></td></tr></table>

```{r fig.height=9, fig.width=16}
latest_date <- tail(names(confirmed_cleaned), 1)
confirmed_latest <- confirmed_cleaned %>% 
  select(`Province/State`, `Country/Region`, !!latest_date) %>%
  rename(Total_Cases = !!latest_date)

ggplot(confirmed_latest, aes(y = reorder(`Province/State`, Total_Cases), x = Total_Cases)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = paste("Total COVID-19 Confirmed Cases by Province", latest_date),
       y = "Province",
       x = "Total Cases (in millions)") +
  theme_wsj() +
  theme(
    title = element_text(size = 24),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold")
    ) +
  scale_x_continuous(labels = function(x) x / 1e6, expand = c(0, 0))
```

- Top3: Ontario, Quebec, Alberta

<table style="width: 900px"><tr><td bgcolor="wheat"><font color="brown" size="5"><b>
&nbsp;&nbsp;3.2 Total COVID-19 Death Cases by Province
</b></font></td></tr></table>

```{r fig.height=9, fig.width=16}
latest_date_death <- tail(names(deaths_cleaned), 1)
deaths_latest <- deaths_cleaned %>% 
  select(`Province/State`, `Country/Region`, !!latest_date_death) %>%
  rename(Total_Deaths = !!latest_date_death)

ggplot(deaths_latest, aes(y = reorder(`Province/State`, Total_Deaths), x = Total_Deaths)) +
  geom_bar(stat = "identity", fill = "darkred") +
  labs(title = paste("Total COVID-19 Death Cases by Province"),
       y = "Province",
       x = "Total Deaths (in thousands)") +
  theme_wsj() +
  theme(
    title = element_text(size = 24),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold")
    ) +
  scale_x_continuous(labels = function(x) x / 1e3, expand = c(0, 0))
```

- Top3: Quebec, Ontario, Alberta

<table style="width: 900px"><tr><td bgcolor="wheat"><font color="brown" size="5"><b>
&nbsp;&nbsp;3.3 COVID-19 Fatality Rate by Province
</b></font></td></tr></table>

```{r fig.height=9, fig.width=16}
combined_data <- left_join(confirmed_latest, deaths_latest, by = c("Province/State", "Country/Region"))
combined_data$Mortality_Rate <- (combined_data$Total_Deaths / combined_data$Total_Cases) * 100
mortality_data <- combined_data %>% 
  select(`Province/State`, Mortality_Rate) %>%
  arrange(desc(Mortality_Rate))

ggplot(mortality_data, aes(y = reorder(`Province/State`, Mortality_Rate), x = Mortality_Rate)) +
  geom_bar(stat = "identity", fill = "darkorange") +
  labs(title = "COVID-19 Mortality Rate by Province",
       y = "Province",
       x = "Mortality Rate (%)") +
  theme_wsj() +
  theme(
    title = element_text(size = 24),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold")
    ) +
  scale_x_continuous(expand = c(0, 0))
```

- Top3: Manitoba, Quebec, British Columbia

<table style="width: 900px"><tr><td bgcolor="wheat"><font color="brown" size="5"><b>
&nbsp;&nbsp;3.4 COVID-19 Infection Rate by Province
</b></font></td></tr></table>

2021 Population Data Source:
https://www12.statcan.gc.ca/census-recensement/2021/dp-pd/prof/details/download-te

```{r fig.height=9, fig.width=16}
population_data <- data.frame(
  "Province/State" = c("Ontario", "Quebec", "British Columbia", "Alberta", "Manitoba", "Saskatchewan", "Nova Scotia", "New Brunswick", "Newfoundland and Labrador", "Prince Edward Island", "Northwest Territories", "Yukon", "Nunavut"),
  Population_2021 = c(14223942, 8501833, 5000879, 4262635, 1342153, 1132505, 969383, 775610, 510550, 154331, 41070, 40232, 36858)
)
names(confirmed_latest)[names(confirmed_latest) == "Province/State"] <- "Province.State"

merged_data <- left_join(confirmed_latest, population_data, by = "Province.State")
merged_data$Infection_Rate <- (merged_data$Total_Cases / merged_data$Population_2021) * 100

ggplot(merged_data, aes(y = reorder(Province.State, Infection_Rate), x = Infection_Rate)) + 
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "COVID-19 Infection Rate by Province",
       y = "Province",
       x = "Infection Rate (%)") +
  theme_wsj() +
  theme(
    title = element_text(size = 24),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold")
  ) +
  scale_x_continuous(expand = c(0, 0))

```

- Top3: Prince Edward Island, Northwest Territories, Quebec

<table style="width: 900px"><tr><td bgcolor="wheat"><font color="brown" size="5"><b>
&nbsp;&nbsp;3.5 Fatality Rate vs Infection Rate by Province
</b></font></td></tr></table>

```{r fig.height=9, fig.width=16}
names(mortality_data)[names(mortality_data) == "Province/State"] <- "Province.State"
final_data <- left_join(merged_data, mortality_data, by = "Province.State")

ggplot(final_data, aes(x = Infection_Rate, y = Mortality_Rate, label = Province.State)) + 
  geom_point(color = "steelblue", size = 10) + 
  geom_text(aes(label = Province.State), vjust=-1, hjust=0.5, size=5) +
  labs(title = "Mortality Rate vs Infection Rate by Province",
       y = "Mortality Rate (%)",
       x = "Infection Rate (%)") +
  theme_wsj() +
  theme(
    title = element_text(size = 24),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold")
  )
```

<table style="width: 900px"><tr><td bgcolor="wheat"><font color="brown" size="5"><b>
&nbsp;&nbsp;3.6 Cluster Analysis
</b></font></td></tr></table>

```{r}
data_for_clustering <- final_data %>% 
  select(Infection_Rate, Mortality_Rate)

scaled_data <- scale(data_for_clustering)
wss <- sapply(1:10, function(k){
  kmeans(scaled_data, centers=k)$tot.withinss
})

plot(1:10, wss, type="b", pch=19, frame=FALSE, 
     xlab="Number of clusters K", 
     ylab="Total within-clusters sum of squares")
```

```{r fig.height=9, fig.width=16}
set.seed(123)
kmeans_result <- kmeans(scaled_data, centers=3)

final_data$cluster <- factor(kmeans_result$cluster)
ggplot(final_data, aes(x=Infection_Rate, y=Mortality_Rate, color=cluster)) + 
  geom_point(size = 10) + 
  geom_text(aes(label=Province.State), vjust=-1.3, hjust=0.5, size=5) +
  labs(title="K-means clustering of Provinces (K=3)",
       x="Infection Rate", y="Mortality Rate") +
  scale_color_manual(values=c("red", "blue", "green")) +
  theme_wsj() +
  theme(
    title = element_text(size = 24),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold")
  )

```

- After cluster analysis, Ontario, Quebec, and British Columbia are grouped together.
- Economic development level is negative related to the mortality rate.

<table style="width: 900px"><tr><td bgcolor="darkred"><font color="white" size="6"><b>
&nbsp;&nbsp;4. Conclusion
</b></font></td></tr></table>

<table style="width: 900px"><tr><td bgcolor="wheat"><font color="brown" size="5"><b>
&nbsp;&nbsp;4.1 Data Bias
</b></font></td></tr></table>

#### 1. Objective bias: The data for Covid_19 depends on the local medical standards. If medical resources cannot keep up, the number of infections might not be recorded, potentially leading to a relative decrease. On the other hand, if there are robust medical resources, the number of infections might relatively increase.

#### 2. Subjective bias: The data for Covid_19 comes from manual counting, and it might be adjusted for various reasons. Therefore, the data from different countries or even different provinces may not necessarily be directly comparable.

<table style="width: 900px"><tr><td bgcolor="wheat"><font color="brown" size="5"><b>
&nbsp;&nbsp;4.2 Further Analysis
</b></font></td></tr></table>

#### This analysis is suitable for various countries to evaluate the performance of their internal provinces or states.

